### 2023-11-28 R/analysis.qmd
minor changes to formatting, mostly added code to define function `prepare_image()`

---
title: "Supplement: Fremont Models"
date: today
format:
  html: 
    embed-resources: true
    number-sections: true
    number-depth: 3
    toc: true
    toc-depth: 3
    smooth-scroll: true
    theme: cosmo
    highlight-style: a11y
execute: 
  eval: true
  echo: true
  warning: false
  error: false
---

```{css}
/*| echo: false */
figcaption {
    display: none !important;
}
```

## Overview

**Goal**: model the distribution of Fremont sites across watersheds.

**Data**: (i) site counts per watershed based on records from the State Historic
Preservation Offices of Utah and Nevada and (ii) environmental data from the
Oregon State University PRISM Group, as well as the USGS.

**Method**: a GAMM with a negative binomial distribution for count data with
dispersion, an offset to account for the area of each watershed, and an
exponential covariance matrix to account for residual spatial autocorrelation.

## R Preamble

```{r libraries}

library(GGally)
library(ggeffects)
library(gt)
library(here)
library(mgcv)
library(patchwork)
library(sf)
library(sfdep)
library(tidyverse)
library(tigris)
library(viridis)

```

Some plotting defaults

```{r}
#| code-fold: true
#| file: "fun-prepare-image.R"

```

## Data

Unit of analysis: watersheds

Dependent variable:\
- a count of **sites** per watershed

Independent variables:\
- **elevation**\
- maize growing degree days (**gdd**, in Celsius)\
- **precipitation** (millimeters)\
- cost-distance (in hours) to **springs** and **streams**\
- **protected** status\
- cost-distance (in hours) to **roads**

The climate variables, precipitation and gdd, were hindcasted for each watershed
using Kyle Bocinsky's [paleocar](https://github.com/bocinsky/paleocar) package
and averaged over the approximately 1,000 year occupational sequence of the
Fremont in the project area. The cost-distance variables are derived from
elevation data using Campbell's hiking function and Djikstra's search algorithm
and averaged over each watershed. Protected status refers to the proportion of
each watershed that is federal land. Note that it and cost-distance to roads are
not explanatory variables, but rather controls on potential sources of sampling
bias, mostly due to taphonomic processes operating on the archaeological
record - basically, modern farms and cities.

Offsets:\
- **area** of each watershed

This is included to account for sampling bias owing to the size of each
watershed. The other two are there to account for sampling bias owing to
variable survey intensity, as well as the potential for human impacts to
cultural resources.

**Note:** not all of these variables make it into the final model as we first
evaluate them for potential colinearity and concurvity. An additional concurvity
test is performed after fitting the final model. Not all of them have smooths
applied in the final model, either.

```{r}

gpkg <- here("data", "western-fremont.gpkg")

watersheds <- read_sf(gpkg, "watersheds")

# the following shapes are just for visualization purposes
utah <- read_sf(gpkg, "utah")

window <- read_sf(gpkg, "window")

roads <- read_sf(gpkg, "roads") |> 
  filter(rttyp %in% c("U", "I")) |> 
  st_intersection(window)

towns <- tigris::places("Utah", progress_bar = FALSE) |> 
  rename_with(tolower) |> 
  select(name) |> 
  filter(name %in% c("Salt Lake City", "Cedar City", "Delta", "Richfield")) |> 
  arrange(name) |> 
  st_centroid()

```

::: {#fig-overview}
![](../figures/overview.jpg){width="100%" fig-alt="overview map of project area"}

Overview map.
:::

### Test for colinearity

```{r}
#| label: fig-correlations
#| code-fold: true
#| out-width: "75%"
#| fig-height: 5
#| fig-asp: 1
#| fig-alt: "Pearson correlation matrix"
#| fig-cap: "Pearson correlation matrix"

diagonal_labels <- function(data, mapping, ...) {

  GGally::ggally_text(
    rlang::as_label(mapping$x),
    col = "black",
    size = 4
  )

}

correlations <- function(data, mapping, color = "black", ...) {

  # get the x and y data to use the other code
  x <- GGally::eval_data_col(data, mapping$x)
  y <- GGally::eval_data_col(data, mapping$y)

  ct <- cor.test(x,y)

  r <- unname(ct$estimate)
  rt <- format(r, digits=2)[1]
  tt <- as.character(rt)

  tt <- stringr::str_c(
    tt, 
    GGally::signif_stars(ct$p.value)
  )

  # plot the cor value
  ggally_text(
   label = tt, 
   mapping = aes(),
   xP = 0.5, 
   yP = 0.5, 
   size = 4,
   color = color,
   ...
  )

}

# there's a bug in ggpairs() where it mislabels the y-axis, so we remove the
# text and ticks from this plot

watersheds |> 
  st_drop_geometry() |> 
  select(elevation, precipitation, gdd, streams, springs, protected, roads) |> 
  ggpairs(
    diag = list(continuous = diagonal_labels),
    lower = list(continuous = wrap("points", alpha = 0.3)),
    upper = list(continuous = correlations)
  ) +
  theme_bw(12) +
  theme(
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )

```

As you can see, there are strong, linear correlations between all the variables,
including the familiar inverse relationship between precipitation and gdd, which
is driven by elevation in the American West. Based on this and a previous
post-hoc concurvity test, we will drop springs.

### Map of covariates

```{r}
#| label: fig-covariates
#| code-fold: true
#| fig-width: 8
#| fig-asp: 0.9
#| fig-alt: "Map of covariates"
#| fig-cap: "Map of covariates"

make_map <- function(x, fill, name) {

  ggplot() +
    geom_sf(
      data = utah,
      fill = "gray95"
    ) +
    geom_sf(
      data = x, 
      aes(fill = {{ fill }}),
      color = "gray90", 
      linewidth = 0.2
    ) +
    scale_fill_viridis(
      name = name,
      option = "mako"
    ) +
    geom_sf(
      data = roads |> filter(rttyp == "I"),
      color = "white",
      linewidth = 0.4
    ) +
    geom_sf(
      data = roads |> filter(rttyp == "U"),
      color = "white",
      linewidth = 0.2
    ) +
    geom_sf(
      data = towns,
      fill = "white",
      color = "gray35",
      shape = 21,
      size = 2
    ) +
    theme_void() +
    theme(
      legend.background = element_rect(fill = "gray95", color = "transparent"),
      legend.justification = c("left", "bottom"),
      legend.position = c(0.65, 0.19),
      legend.title = element_text(size = rel(0.9))
    )

}

gdd <- make_map(
  watersheds,
  fill = gdd,
  name = "GDD (°C)"
)

ppt <- make_map(
  watersheds,
  fill = precipitation,
  name = "PPT (mm)"
)

streams <- make_map(
  watersheds,
  fill = streams,
  name = "Streams\n(hours)"
)

protected <- make_map(
  watersheds,
  fill = protected, 
  name = "Protected"
)

bb8 <- st_union(watersheds, utah) |> st_bbox()

dy <- bb8[["ymax"]] - bb8[["ymin"]]
dx <- bb8[["xmax"]] - bb8[["xmin"]]

fn <- here("figures", "covariates.jpg")

ggsave(
  plot = (gdd + ppt) / (streams + protected),
  filename = fn,
  width = 5.75,
  height = 5.75 * (dy/dx),
  dpi = 600
)

prepare_image(fn)

springs <- make_map(
  watersheds,
  fill = springs, 
  name = "Springs\n(hours)"
)

routes <- make_map(
  watersheds,
  fill = roads, 
  name = "Roads\n(hours)"
)

all_covariates <- (gdd + ppt + streams) / (springs + protected + routes)

remove(gdd, ppt, streams, springs, routes, protected)

all_covariates

```

### Bocinsky 2014 Thresholds

```{r}
#| label: fig-thresholds
#| code-fold: true
#| fig-width: 8
#| fig-asp: 0.45
#| fig-alt: "Histograms of temperature and precipitation"
#| fig-cap: "Histograms of temperature and precipitation"

thresholds <- tibble(
  xintercept = c(300, 1000),
  x = c(320, 1020),
  y = 72,
  variable = c("precipitation", "gdd"),
  label = c("Minimum threshold\n300 mm", "Minimum threshold\n1000°C GDD")
)

values <- watersheds |>
  st_drop_geometry() |> 
  select(precipitation, gdd) |> 
  pivot_longer(
    everything(),
    names_to = "variable"
  )

pretty_breaks <- function(x) {

  xmin <- round(min(x), -2)
  xmax <- round(max(x), -2)

  seq(xmin, xmax, by = 200)

}

ggplot() +
  geom_histogram(
    data = values,
    aes(value),
    color = "white",
    binwidth = 100,
    center = 50
  ) +
  geom_vline(
    data = thresholds,
    aes(xintercept = xintercept),
    linewidth = 1.2,
    color = "darkred"
  ) +
  geom_text(
    data = thresholds,
    aes(x, y, label = label),
    color = "darkred",
    hjust = 0,
    vjust = 1
  ) +
  facet_wrap(
    ~variable, 
    scale = "free_x",
    labeller = labeller(
      variable = c(precipitation = "Precipitation (mm)", 
                   gdd = "Maize GDD (°C)")
    ),
    strip.position = "bottom"
  ) +
  scale_x_continuous(breaks = pretty_breaks) +
  labs(
    x = NULL,
    y = "Count of watersheds"
  ) +
  theme_bw(12) +
  theme(
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = rel(1)),
    strip.placement = "outside"
  )

```

Here's what that threshold looks like in geographic space.

```{r}

watersheds <- watersheds |> 
  mutate(
    in_niche = ifelse(gdd >= 1000 & precipitation >= 300, "Yes", "No"),
    in_niche = factor(in_niche)
  )

```

```{r}
#| label: fig-thresholds-map
#| code-fold: true
#| fig-width: 4
#| out-width: "4in"
#| fig-alt: "Map of maize niche thresholds"
#| fig-cap: "Map of maize niche thresholds"

thm <- ggplot() +
  geom_sf(
    data = utah,
    fill = "gray95"
  ) +
  geom_sf(
    data = watersheds, 
    aes(fill = in_niche),
    color = "gray90", 
    linewidth = 0.2
  ) +
  scale_fill_viridis(
    name = "In Niche?",
    option = "mako",
    end = 0.75,
    discrete = TRUE
  ) +
  geom_sf(
    data = roads |> filter(rttyp == "I"),
    color = "white",
    linewidth = 0.4
  ) +
  geom_sf(
    data = roads |> filter(rttyp == "U"),
    color = "white",
    linewidth = 0.2
  ) +
  geom_sf(
    data = towns,
    fill = "white",
    color = "gray35",
    shape = 21,
    size = 2
  ) +
  theme_void() +
  theme(
    legend.background = element_rect(fill = "gray95", color = "transparent"),
    legend.justification = c("left", "top"),
    legend.position = c(0.66, 0.57)
  )

thm

```

Is there a significant difference in elevation between those in the niche and
those outside it?

```{r}

with(watersheds, t.test(elevation ~ in_niche))

```

## Analysis

Here, we go through iterations of model fitting and evaluation, with changes to
correct for problems like autocorrelation and concurvity (the smoothed version
of co-linearity in the predictors). For each fitted model, we provide a model
summary, diagnostic plots, and visualizations of partial effects (sometimes
called marginal responses) for each smooth term.

```{r}

fm <- sites ~ 
  s(precipitation, k=5) +
  s(gdd, k=5) +
  s(streams, k=5) +
  s(protected, k=5) + 
  offset(log(area))

fremont <- gam(
  fm, 
  data = watersheds,
  family = poisson, 
  select = TRUE
)

summary(fremont)

```

### Dispersion

First, an h-test for scaled dispersion in a poisson model. In a poisson model,
the variance is supposed to equal the mean. If this holds, then $\alpha$ in
$Var(y) = E(y) + \alpha \cdot E(y)$ should equal zero. A simple, intercept-only
linear model can test this idea.

```{r}

# code adopted from AER::dispersiontest()

observed <- model.response(model.frame(fremont))
estimate <- fitted(fremont)

variance <- ((observed - estimate)^2 - observed)/estimate

dispersion_model <- lm(variance ~ 1)

```

```{r}
#| code-fold: true

tibble(
  estimate = as.vector(coefficients(dispersion_model)[1]),
  statistic = as.vector(summary(dispersion_model)$coef[1, 3]),
  null = 0,
  p.value = pnorm(statistic, lower.tail = FALSE)
) |> 
  mutate(across(everything(), round, digits = 4)) |> 
  gt(
    rowname_col = NULL
  ) |> 
  tab_caption(
    caption = gt::html("<p>t-test for Overdispersion<br>Alternative: &alpha; > 0</p>")
  ) |> 
  tab_options(
    table.align = "left",
    container.width = pct(35)
  )

```

Clearly, there is over-dispersion in this model. To correct for this, we re-run
the model with a `negative binomial` error distribution, letting the model
adjust to $\alpha$. We choose a negative binomial over a quasipoisson as it uses
maximum likelihood.

```{r}

fremont <- gam(
  fm, 
  data = watersheds,
  family = nb, 
  select = TRUE
)

summary(fremont)

```

### Residual Spatial Autocorrelation

Monte Carlo simulations of Moran's I to test for spatial autocorrelation in the
residuals (using sfdep).

```{r}

neighbors <- st_contiguity(watersheds)
weights <- st_weights(neighbors)

global_moran_perm(
  residuals(fremont),
  neighbors,
  weights
)

```

There's significant spatial autocorrelation. To fix this, we will use a
generalized additive mixed-model (GAMM) with an exponential spatial correlation
structure. This is not strictly correct, as it assumes spatial continuity of
points, not polygons, but it's a useful first approximation.

### Spatial Correlation Structure

Add coordinates for watershed centroids to the dataset.

```{r}

xy <- watersheds |> 
  st_centroid() |> 
  st_coordinates() |> 
  as_tibble() |> 
  rename_with(tolower)

watersheds <- bind_cols(watersheds, xy)

```

```{r}

fremont <- gamm(
  fm,
  correlation = corExp(form = ~x+y),
  data = watersheds,
  family = nb,
  verbosePQL = FALSE
)

summary(fremont$lme)

summary(fremont$gam)

```

By accounting for spatial autocorrelation, the expected degrees of freedom for
gdd and cost-distance to streams are now both 1, which suggests that there are
no non-linear effects for those covariates.

### Concurvity

A post-hoc test of non-linear correlations in the predictors.

```{r}

concurvity(fremont$gam)

```

Not good...

### Final model

Cost-distance to roads is non-significant, so we'll remove that variable from
the model entirely.

```{r}

fm <- sites ~ 
  s(precipitation, k=5) +
  gdd +
  streams +
  protected +
  offset(log(area))

fremont <- gamm(
  fm,
  correlation = corExp(form = ~x+y),
  data = watersheds,
  family = nb,
  verbosePQL = FALSE
)

summary(fremont$lme)

summary(fremont$gam)

```

### Final Moran's I

One more Moran's I test to see if we really got a hold of the spatial
autocorrelation. We'll do this with the normalized residuals of the lme inside
the GAMM as that incorporates the exponential covariance matrix.

```{r}

global_moran_perm(
  residuals(fremont$lme, type = "normalized"),
  neighbors,
  weights
)

```

Looks good!

### Variance Inflation Factor

There can't be concurvity in the model anymore, as there is only one smoothed
term. However, we should now check for potential multi-collinearity in the
parameteric terms.

```{r}

# borrowing this code from mgcv.helper
# https://github.com/samclifford/mgcv.helper/blob/master/R/vif.gam.R

vif <- function(object){

  obj.sum <- mgcv::summary.gam(object)

  # estimate of standard deviation of residuals
  s2 <- object$sig2

  # data used to fit the model
  X <- object$model

  # n observations
  n <- nrow(X)

  # omit the intercept term, it can't inflate variance
  v <- -1

  # variance in estimates
  varbeta <- obj.sum$p.table[v,2]^2

  selected_col <- row.names(obj.sum$p.table)[v]
  selected_col <- gsub("TRUE", "", selected_col)

  # variance of all the explanatory variables
  varXj <- apply(X=X[, selected_col],MARGIN=2, var)

  # the variance inflation factor, obtained by rearranging
  # var(beta_j) = s^2/(n-1) * 1/var(X_j) * VIF_j
  VIF <- varbeta/(s2/(n-1)*1/varXj)

  tibble::tibble(
    variable = names(VIF),
    vif = VIF
  )

}

vif(fremont$gam)

```

The VIF for gdd is a little high, but within acceptable limits.

## Results

### Marginal response

Here we estimate the marginal responses (also known as partial effects) for each
covariate by letting the target covariate vary and holding the other covariates
at their means and predicting the site count (the response). Here we use the
{ggeffects} package.

```{r}

margins <- ggpredict(fremont, condition = list(area = mean(watersheds$area))) |>
  unclass() |>
  bind_rows() |>
  as_tibble() |> # the "ggeffects" class doesn't want to let go
  filter(group != "area") |>
  rename(
    "y" = predicted,
    "ymin" = conf.low,
    "ymax" = conf.high,
    "variable" = group
  )

```

```{r}
#| label: fig-partials
#| code-fold: true
#| fig-height: 6
#| fig.asp: 0.93
#| out-width: "70%"
#| fig-alt: "Partial dependence plots"
#| fig-cap: "Partial dependence plots"

margins <- margins |> 
  mutate(
    # hacky way to "clip" the ribbon before expanding in ggplot
    ymax = ifelse(ymax > 100, 100, ymax)
  )

response_labels <- tibble(
  variable = c("precipitation", "gdd", "streams", "protected"),
  label = c("Precipitation\n(mm)", 
            "Maize GDD\n(°C)", 
            "Streams\n(hours)", 
            "Protected\n(Federal Land)"),
  x = with(watersheds, c(max(precipitation), min(gdd), max(streams), min(protected))),
  y = max(margins$ymax),
  hjust = c(0.95,0.05,0.95,0.05)
)

scale_pretty <- function(x, n = 4L) {

  brks <- pretty(x, n = n)

  nbrks <- length(brks)

  i <- ifelse(
    nbrks > n,
    c(1, nbrks),
    nbrks
  )

  brks[-i]

}

marginal_response <- ggplot(margins, aes(x, y)) +
  geom_ribbon(
    aes(ymin = ymin, ymax = ymax, fill = variable)
  ) + 
  scale_fill_manual(
    values = c("#4F3824", "#D1603D", "#DDB967", "#848FA5")
  ) +
  geom_line(
    linewidth = 2,
    color = alpha('white', 0.75),
    lineend = "round"
  ) +
  geom_line(
    linewidth = 1,
    lineend = "round"
  ) +
  geom_label(
    data = response_labels,
    aes(x, y, label = label),
    size = 4,
    hjust = response_labels$hjust,
    vjust = 1,
    label.size = NA,
    alpha = 0.5,
    lineheight = 0.88
  ) +
  labs(
    x = NULL,
    y = "Number of sites"
  ) +
  facet_wrap(
    ~variable, 
    scale = "free_x"
  ) +
  scale_x_continuous(breaks = scale_pretty) +
  coord_cartesian(ylim = c(0, 100)) +
  theme_bw(12) +
  theme(
    axis.title.y = element_text(size = 13),
    legend.position = "none",
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank()
  )

fn <- here("figures", "marginal_response.jpg")

ggsave(
  plot = marginal_response,
  filename = fn,
  width = 5,
  height = 5 * 0.93,
  dpi = 600
)

prepare_image(fn)

marginal_response

```

### Map

```{r}

watersheds <- watersheds |> 
  mutate(
    estimate = fitted(fremont$gam) |> as.vector(),
    residuals = residuals(fremont$lme, type = "normalized")
  )

```

```{r}
#| label: fig-model-map
#| code-fold: true
#| fig-width: 8
#| fig-asp: 0.9
#| fig-alt: "Map of estimated density"
#| fig-cap: "Map of estimated density"

obs <- make_map(
  watersheds,
  fill = ifelse(sites == 0, log(0.00001), log(sites/area)), 
  name = "Observed\nLog Density"
)

est <- make_map(
  watersheds,
  log(estimate/area), 
  "Estimated\nLog Density"
)

res <- make_map(
  watersheds,
  fill = residuals, 
  name = "Normalized\nResiduals"
)

fn <- here("figures", "model-map.jpg")

ggsave(
  plot = (obs + est),
  filename = fn,
  width = 5.75,
  height = 5.75 * (dy/(2*dx)),
  dpi = 600
)

prepare_image(fn)

fn <- here("figures", "threshold_map.jpg")

ggsave(
  plot = (est + thm),
  filename = fn,
  width = 5.75,
  height = 5.75 * (dy/(2*dx)),
  dpi = 600
)

prepare_image(fn)

(obs + est) / (res + thm)

```

Please note that these are normalized residuals from the LME model.

```{r}
#| echo: false

# save model 
saveRDS(fremont, file = here("data", "western-fremont-model.Rds"))

```

## Session Info

```{r}
#| code-fold: true

# save the session info as an object
pkg_sesh <- sessioninfo::session_info(pkgs = "attached")

# inject the quarto info
pkg_sesh$platform$quarto <- paste(
  quarto::quarto_version(), 
  "@", 
  quarto::quarto_path()
  )

# print it out
pkg_sesh

```