### 2024-01-31 manuscript/cover-letter-response-to-reviews.md
draft response to reviewers

We thank the reviewers for their many insightful comments and criticisms
regarding our paper.

All minor comments about grammatical errors, poor wording, and the like have
been addressed in the text. Major comments from reviewers are listed here with
comments about how we addressed them in the text. Reviewer two had the most
substantial comments and criticisms, so most of this involves responses to them.

### Reviewer #1

1.  Description of Fremont extent is misleading and vague.

    Response: We updated the description of the Fremont extent in the background
    section to make it clearer that there is evidence for Fremont foragers in ID
    and WY and Fremont farmers in NV and CO, rather than leaving it vague
    whether they might have all been farmers.

2.  Description of project area is inaccurate.

    Response: We updated this section to fix inaccurate descriptions. We also
    updated the overview map, adding major roads and some important nearby
    cities to give the reader more landmarks for orientation.

3.  The impossibility of dry farming does not necessarily entail irrigation.

    Response: This is correct, so we have added text to the discussion to
    clarify. In the subsection title, we also replaced 'irrigation' with 'water
    management', and in the text, we change "making irrigation ... necessary" to
    "making irrigation or some other form of water management ... necessary."

### Reviewer #2

1.  Why not high precipitation?

    Response: Our answer to this involves the trade-off with temperature in the
    project area. As elevation increases, precipitation increases but
    temperature decreases. So, the places with the most precipitation are also
    the ones most likely to experience frost and to have shorter growing
    seasons. This means the best place for maize is not necessarily the place
    with the most precipitation. We've updated the discussion, rearranging some
    comments into a new section entitled "Trade-off Between Precipitation and
    Temperature," to emphasize this point.

2.  The limits of rain-fed maize farming should be discussed more explicitly.

    Response: Our analysis of the Fremont maize farming niche is suggestive of
    its limits, but defining the true limits requires actualistic research,
    hence our reluctance to offer specific numbers on this subject. We do,
    however, discuss the results of some actualistic research, particularly the
    work referenced by Bocinsky in his 2014, and compare our results to that.
    This is the basis for our claim that the Fremont likely lived outside of the
    rain-fed maize farming niche, thus requiring some kind of water management
    on their part.

3.  Details of data collection are too vague.

    Response: We updated the Site Data section to more accurately represent how
    we came by these data.

    -   The NSF project alluded to in the previous version of this paper now
        includes its award number along with several citations to research
        coming out of that project.
    -   The "stratified random sample" comment was a bit misleading. That was
        done for a portion of the sites (we sampled a fraction of site forms in
        each county), but not systematically to produce the entire sample, which
        was mostly arrived at in a more opportunistic manner.
    -   To get room counts, we used the list of Fremont villages found in
        Coltrain and Leavitt (2002), and cross-referenced those with the
        original site forms, excavation reports, and Adrien Mooney's 2014 BYU MA
        thesis, which tries to reconstruct some poorly reported excavations. For
        many of these sites, the estimates can vary wildly, so we erred on the
        side of choosing the more conservative estimates.
    -   In spite of these limitations with our data collection scheme, we note
        that the current dataset represents one of the largest Fremont samples
        used for this sort of analysis to date.

4.  What kind of sites are these?

    Response: The reviewer requests additional information about site types, but
    we do not currently have the data to answer that question. Mainly, our data
    collection efforts were focused on identifying as many Fremont related sites
    as possible. In fact, we have done this now for a considerable portion of
    the state of Utah, though efforts are ongoing. We are also trying to go back
    and record as much attribute data as is available and will hopefully be able
    to report on these efforts in the near future.

5.  How do we know that these are farming and not foraging adaptations?

    Response: The reviewer notes that site types are needed to differentiate
    hunting sites from farming sites, and thus to determine whether settlement
    decisions were shaped by constraints on farming vs constraints on foraging.
    This is a really important point that goes right to the heart of our model
    interpretation. If the sites in our analysis are not primarily associated
    with farmers, then we cannot use them to understand the maize farming niche.
    Unfortunately, there's just not a lot of good data to really answer this
    question. The available stable isotope data that exists for the Fremont
    suggests that they were, for the most part, consuming maize at a rate
    comparable to maize farmers in the San Juan Basin, including Chaco Canyon.
    See, for example, Coltrain and Leavitt 2002 and Coltrain 1993. As Coltrain
    and Leavitt point out, the one exception to this general pattern is the
    Fremont around Bear River and the Willard Bay wetlands, but even there
    evidence of a largely hunting based diet is limited to the last few
    centuries in their sequence (around 900 CE their subsistence pattern becomes
    more mixed, and only around 1100 do they fully adopt foraging). This is not
    to say that foraging wasn't important to the Fremont. Just that farming was
    likely the critical constraint on settlement for most Fremont for most of
    their history (Yaworsky 2023; Spangler 2000, 2002).

6.  The Ideal Free Distribution section is overly detailed.

    Response: The reviewer thinks the discussion of the IFD model is overly
    detailed because "the analysis doesn't rank habitat suitability or site
    association with suitability." The reviewer suggests a simpler discussion
    that contextualizes the analysis "within the literature on SDMs." In
    response, we would like to note three things.

    -   First, we agree that the section was overly detailed. In fact, the main
        logic we wanted to convey is already in the intro, so the section is not
        just overly detailed but basically unnecessary, so we have removed it
        entirely.

    -   Second, distressingly, it's precisely the goal of this paper to estimate
        relative differences in habitat suitability across watersheds and
        environmental gradients, but the reviewer's comment suggests that we did
        not do a good job of making this clear. So, we have updated the
        introduction (particularly, the last paragraph) and the discussion to
        make this goal more explicit.

    -   Third, as the reviewer requested, we also explicitly connected this to
        the SDM literature by noting that ecologists will sometimes refer to
        SDMs as "Habitat Suitability Models" (e.g., Rowden et al 2017; Kort et
        al 2020; Bowden et al 2021).

7.  Goldilocks Zone concept is misleading.

    Response: The title of the subsection was changed to "Trade-off Between
    Precipitation and Temperature." The expression is now described explicitly
    as a "metaphor," and it's made clear that the reader should not
    over-interpret it for the reasons listed by the reviewer. We have also
    removed all but a single reference to it.

### Reviewer #4

1.  There are several reference errors and omissions.

    Response: These have been corrected. The more recent references to Fremont
    scholarship offered by the reviewer have also been added.

2.  But are they optimizers?

    Response: This paper is not meant to be a defense of the optimality
    assumption, just an example of its application, so we would prefer not to
    litigate that question here. We simply propose that Fremont farmers located
    themselves in places that yield the highest probabilities of a successful
    harvest. This assumption only requires that the Fremont had an intimate
    knowledge of their local ecology and leveraged this knowledge to make
    intelligent decisions regarding the placement of farming sites. Some of the
    wording in the introduction was changed to emphasize this.

3.  Trade-offs between spatial and temporal resolution?

    Response: We added an entire section to the discussion on this specific
    topic. It has the title "The Data Quality Problem in Fremont Research."
    Hopefully, it helps to assuage the reviewer's concern about a lack of
    temporal resolution in our data, which we fully admit is an issue. While
    having precise chronological control would be preferable, we argue that our
    results are not meaningless because "typically better watersheds," meaning
    watersheds that were warmer and wetter for most of the Fremont sequence,
    "would have attracted maize farmers more often and, over time, would have
    accumulated a larger archaeological maize farming assemblage, too." So, the
    results aren't perfect, but they are at least suggestive of the real drivers
    of Fremont residential choice behavior.