### 2024-01-31 manuscript/manuscript.qmd
all changes made to meet reviewer concerns

---
title: The Fremont Frontier
subtitle: Living at the Margins of Maize Farming
date: today
author:
  - name: Kenneth B. Vernon
    email: Kenneth.Vernon@colorado.edu
    affiliations: 
      - id: colorado
        address: | 
          Center for Collaborative Synthesis in Archaeology,  
          Institute of Behavioral Science,  
          University of Colorado, Boulder,  
          1440 15th Street,  
          Boulder, CO 80302
    attributes:
        corresponding: true 
  - name: Peter M. Yaworsky
    affiliations:
      - id: aarhus-culture
        address: |
          Department of Culture and Heritage Studies, 
          School of Culture and Society, 
          Aarhus University, 
          Moesgård Allé 20, building 4216, 
          8270 Højbjerg, Denmark   
      - id: aarhus-biodiversity
        address: |
          Center for Biodiversity Dynamics in a Changing World, 
          Department of Biology, 
          Aarhus University, 
          Ny Munkegade 114-116, 
          8000 Aarhus C, Denmark
  - name: Weston C. McCool
    affiliations:
      - id: utah-anthropology
        address: | 
          University of Utah Department of Anthropology, 
          260 S. Central Campus Drive, RM 4625, 
          Salt Lake City, UT 84112
  - name: Jerry D. Spangler
    affiliations:
      - ref: utah-anthropology
  - name: Simon Brewer
    affiliations:
      - id: utah-geography
        address: |
          University of Utah Department of Geography, 
          260 S. Central Campus Drive, RM 4625, 
          Salt Lake City, UT 84112
      - id: gcsc
        address: |
          University of Utah Global Change and Sustainability Center, 
          115 South 1460 East, RM 234 FASB, 
          Salt Lake City, UT 84112
  - name: Brian F. Codding
    affiliations:
      - ref: utah-anthropology
      - ref: gcsc
abstract: |
  The Fremont provide an important case study for examining the resilience of ancient farmers to climatic downturns, for they lived at the far northern margin of intensive maize agriculture in the American West, where the constraints on maize production are made abundantly clear. Using a tree-ring and simulation-based reconstruction of average annual precipitation and temperature, along with cost-distance to perennial streams, we model spatial variability in Fremont site density in the eastern Great Basin. The results of our analysis have implications for defining the ecological envelope in which farming is a viable strategy across this arid region and can be used to predict where and why maize farming strategies might evolve and eventually collapse as climate changes over time.
keywords:
  - Ideal Free Distribution
  - Great Basin
  - Paleoclimate Reconstruction
corresponding_author: Kenneth.Vernon@colorado.edu
linenumbers: false
bibliography: bibliography.bib
format: 
  simple-pdf: default
  docx: default
execute: 
  echo: false
  error: false
  warning: false
---

# Introduction

Archaeological populations of subsistence maize farmers known collectively as
"the Fremont" lived at the far northern periphery of maize farming in western
North America, in an area encompassing much of the modern state of Utah north of
the Virgin and Colorado Rivers, from roughly 2000 to 700 years BP [@madsen1989;
@madsen_simms1998]. Conditions in this area are extremely harsh and inhospitable
to maize farming, so the fact that the Fremont not only pulled it off but also
thrived while doing so poses an interesting puzzle. How did they do it?
Archaeologists interested in this question have typically focused on strategies
directly related to farming and subsistence, like irrigation methods, trade, and
seasonal high elevation hunting [e.g., @barlow_etal2008; @boomgarden_etal2019;
@janetski2002; @hart_etal2021; @madsen_simms1998; @metcalfe_larrabee1985;
@patterson_flanigan2010; @spangler1993; @morgan_etal2012]. Here, however, we
come at the problem from a slightly different angle. We seek to understand not
so much *what* they did to make farming effective, but *where* they chose to do
it, and why they chose those places over others [e.g., @bocinsky_kohler2014;
@thomson_macdonald2020]. Those are not entirely separate questions, of course,
as they trade-off each other, but we think a geographic analysis can actually
help illuminate some of the costs and benefits of the other strategies adopted
by the Fremont. A geographic approach will also help us to investigate the ways
that climate change might have structured those costs and benefits, and by
extension the settlement decisions of Fremont farmers.

As a test case, we focus on those Fremont living in the eastern Great Basin of
western Utah, which we refer to as the western Fremont (see @fig-overview). The
extreme desert conditions that the western Fremont adapted to provide a useful
backdrop for exploring Fremont settlement, for variation in that environment
exhibits abrupt and quite dramatic changes over short distances of space and
time, thus heightening the differences between areas the western Fremont did and
did not occupy. We explore those differences from a socio-ecological
perspective, focusing on dynamic interactions between competing individuals on
the one hand and individuals and their environment on the other
[@bird_oconnell2006; @kennett_etal2006]. This dynamic is captured nicely by the
Ideal Free Distribution model from Behavioral Ecology, as it represents
individual decisions about where to live as a choice between habitats whose
benefits to the individual may be constrained by what other individuals are
doing. The technical term for these benefits is 'suitability', which the IFD
defines as some function of both the habitat's intrinsic or pristine
environmental quality and its current population size [@fretwell_lucas1969;
@codding_bird2015; @winterhalder_etal2010].

The fundamental assumption underlying the IFD is that individuals will behave
optimally, that they will choose to settle the habitat with the highest
suitability first, meaning the one that offers them the greatest ratio of
benefits to costs relative to the available alternatives. Of course, this comes
with the usual caveats. The claim is not that individuals are infallible, that
they will always do exactly the right thing, but that their choices will tend to
approximate the best strategy given their constraints and trade-offs.

Assuming the western Fremont optimized their agricultural practices, we can use
an empirical or inductive species distribution model [SDM, @elith_etal2006;
@elith_leathwick2009] of Fremont sites to estimate relative differences in maize
suitability across specific environmental gradients and geographic locations
[@jochim2022; @vernon_etal2022a; @yaworsky_etal2020; @elith_leathwick2009]. In
ecology, you will sometimes hear this approach referred to as "ecological niche
modeling" [e.g., @feng_etal2019; @sillero_barbosa2021; @sillero_etal2021] or
even more explicitly as "habitat suitability modeling" [e.g., @bowden_etal2021;
@rowden_etal2017; @dekort_etal2020]. The basic idea here is that we can look at
the distribution of Fremont sites as a reliable indicator of the sorts of
available environmental conditions that would best promote maize farming in an
arid landscape. To put that in more blatantly economic terms, we are assuming
that their environmental preferences are shaped primarily by the constraints of
maize farming and revealed in their residential choice behavior. This allows us
to build a tentative model of the ecological niche for maize farming and to make
some defeasible inferences about its limiting conditions.

::: {#fig-overview fig-align="center"}
![](../figures/overview.jpg){width="5in"}

On the left is an overview map showing the project area with modern satellite
imagery, so the green areas indicate plant growth, the beige areas, empty
desert. The green areas are also the mountainous areas. Thick orange lines are
interstate highways. Thin orange lines are major state highways. For
visualization purposes, the map on the right shows the log-transformed density
of feature-weighted archaeological sites. If no sites have been recorded in a
watershed, log(1e-5) is shown.
:::

# Background

## The Fremont

Archaeologists have nominated a number of artifact types and artifact properties
as candidate traits to distinguish the Fremont from their neighbors in space and
time, including metates with secondary grinding surfaces, large
semi-subterranean pithouses, various clay figurines, trapezoidal rock art,
one-rod-and-bundle basketry, painted white and corrugated gray ceramics, and
elongated corner-notched arrow points [@madsen_simms1998]. None of these,
however, nor any combination of them, has yet proven up to the task of bringing
these diverse people under a single, all-encompassing definition
[@madsen_simms1998]. Nevertheless, the Fremont do stand out, especially at their
population apex roughly 1000 years BP.

Around this time, the Fremont likely reached their greatest geographic extent,
inhabiting an area that encompassed most of the modern state of Utah
[@janetski_etal2011], with varying levels of support for Fremont foragers in
Idaho [@dean1992prehistoric] and Wyoming [@hakiel1987archery; @smith1992fremont]
and Fremont farmers in Nevada [@hockett1998; @cole2012a] and Colorado
[@baker1999]. These modern political boundaries roughly overlap with the eastern
Great Basin and Colorado Plateau physiographic provinces.

::: {#fig-elevation fig-align="center"}
![](../figures/elevation-everything.jpg){width="4in"}

Linear responses of the centered and scaled average water-year precipitation
(mm), average growing-season maize gdd (C), and cost-distance to streams (hours)
to changes in elevation (m). See methods for how these values were calculated
and results for more details.
:::

A basin and range topography dominates the landscape of the Great Basin, with
large, dry valleys and endorheic basins punctuated by north-south trending
mountain ranges. Changes in elevation in this setting are quite dramatic, on the
order of several thousand meters over a short east-west transect. As elevation
drives virtually every climatological process in the region, the local ecology
is characterized by equally dramatic extremes [@billings1951; @grayson2011]. As
shown in @fig-elevation, net water-year precipitation in the eastern Great Basin
tends to increase with elevation, while maize growing degree days (GDD) over the
growing-season tends to decrease. An important trade-off thus exists between
temperature and precipitation, with warmer and drier conditions at lower
elevations and cooler and wetter conditions at higher elevations.

These are the environments into which the Fremont first emerged in western Utah.
Models for describing their origins, while diverse, tend to fall into the
familiar categories of migration and diffusion. Where models of the former
migration variety insist that Ancestral Puebloan farmers migrated northward from
the US Southwest, bringing with them drought-tolerant maize varieties and maize
farming [@madsen_berry1975; @kidder1924], models of the latter diffusion variety
suggest that the Fremont first emerged within these areas as local Archaic
populations transitioned away from mobile foraging, adopting semi-sedentary
agricultural practices from the Southwest in a piecemeal fashion over several
hundred years [@jennings1978; @winter_wylie1974]. While archaeologists have
marshaled an impressive array of arguments and counter-arguments for these two
classes of models, a general consensus appears to be coalescing around the idea
that the truth lies somewhere in between, that the interaction between a
lingering Archaic population existing at extremely low population levels and a
burgeoning community of migrant Southwest farmers together produced the complex
we now call the Fremont [@patterson2015; @spangler1993; @spangler2000;
@spangler2013; @searcy_talbot2015; @simms2008].

Models that try to explain *why* Fremont maize farmers emerged in the far
northern periphery of the US Southwest tend to agree - more or less - that at
some point farming became a more economical alternative to foraging in the
region. Where they differ, it is in whether things got better for farming
starting roughly 2000 years BP [@benson2011; @coltrain_leavitt2002;
@matson_etal1988] or worse for foraging [@barlow2002; @broughton_etal2010;
@cannon2000; @cannon2001]. In support of the *farming-got-better* model,
archaeologists draw on evidence for a suite of favorable climatic conditions
that came together around 1600 years BP, including warmer temperatures and
increased precipitation (both winter and summer), as well as the expansion of
grasslands [@grayson2006; @hemphill_wigand1995; @madsen2000]. The greatest
expansion of the Fremont, in fact, occurred around 1000 years BP, around the
time the Medieval Climate Anomaly (MCA) introduced warmer summer temperatures
that would have increased the intensity of summer monsoons in the Southwest,
pushing them further north into the Great Basin and onto the Colorado Plateau
[@grayson2006], a pattern reflected in numerous tree ring records [e.g.,
@graybill1990; @leavitt1994; @stine1990; @salzer_etal2014], though some
complications do exist for this suggested pattern [@hart_etal2021].

In response to these sorts of arguments, those who favor the
*foraging-got-worse* model [@hart_etal2021; @barlow2002] argue that conditions
promoting maize farming would also tend to make foraging a more productive
strategy. Plus, farming is an extremely costly endeavor, especially when
compared to foraging, including lots of upfront investment of time and energy,
so improving conditions favorable to maize agriculture are not by themselves
sufficient to explain its origins. To fill that lacuna, archaeologists point to
accumulating evidence suggesting that the Fremont transition to agriculture
accompanied periods of sustained population growth [@barlow_etal2008;
@codding_etal2022; @spangler2013] that likely resulted in the depressed
availability of wild resources [@barlow2002; @janetski1997; @simms1986]. So,
demographic pressure reduced the efficiency of foraging strategies, thus making
maize agriculture a more profitable alternative. And that, in turn, led the
Fremont to adopt farming as their primary mode of subsistence.

Once they entered the farming niche, the Fremont quickly developed several
adaptations for coping with the marginal and stochastic conditions in the
eastern Great Basin and the Colorado Plateau, the most important almost
certainly being irrigation (for more on this, see the discussion). On the
Colorado Plateau, the eastern Fremont settled deep within the steep canyons and
washes that flanked the tributaries of major rivers like the Green and Yampa
[@yaworsky2021], in places like Nine Mile Canyon [@spangler1993] and Range Creek
Canyon [@hart_etal2021]. In the eastern Great Basin, the western Fremont chose
to settle on open, alluvial fans and stream terraces along the lower slopes of
most ranges [@janetski_talbot2000; @madsen_simms1998; @simms2008], at places
like Nephi Mounds [@sharrock_marwitt1967], Pharo Village [@marwitt1968], Median
Village [@marwitt1970], Evans Mound [@berry1972; @dodd1982], and Snake Rock
Village [@aikens1967].

Given the extreme aridity of the west, they were still susceptible to drought,
of course, even in those more suitable areas, with a single short growing season
making agricultural returns highly variable. To cope with these uncertainties,
Fremont farmers may have adopted additional risk-mitigation strategies like
spreading farm plots among different micro-environments, a technique otherwise
known as plot diversification [@spangler1993; @spangler2013; @patterson2015].
The Fremont also relied heavily on crop storage both to provide a surplus during
the winter season and to make-up for poor yields [@spangler_etal2019].

# Materials and Methods

## The Project Area

Because we are focusing on the western Fremont, we constrain the current project
area to the eastern Great Basin in western Utah, as shown in @fig-overview,
where the Great Basin is defined in the hydrologic sense using watershed
boundaries [@grayson2011]. On the far eastern edge of this region, a series of
north-south ranges, including the Wasatch Mountains in the north and the Sevier
Plateau in the south, together stand as the primary physiographic boundaries
separating the majority of the Great Basin from the Colorado Plateau. On its far
western edge, a series of north-south ranges straddle the Nevada state line or
lie just inside it, including the Snake Range where Great Basin National Park is
located. Its northern periphery includes the Great Salt Lake (GSL) and the West
Desert, its southern boundary the Bull and Pine Valley Mountains that separate
the Great Basin from the Virgin River watershed just north of St. George, UT.
This is a region approximately 107,000 km$^2$ in area, encompassing virtually
all of the Bonneville Basin in western Utah.

These boundaries also serve as geographic borders between the western Fremont
and other contemporaneous populations. Around the GSL and the Utah-Nevada
border, Fremont farming gives way to Archaic populations engaged in intensive
foraging modes of subsistence. The Wasatch Range and Tushar Mountains separate
the western from the eastern Fremont on the Colorado Plateau, and the Pine
Valley Mountains serve as a border between the Fremont and the Virgin Ancestral
Puebloan to the south [@simms2008].

## The Unit of Analysis

The unit of analysis for this research is the HUC10 watershed (n = 183), as
defined by the Watershed Boundary Dataset (WBD) developed by the US Geological
Survey and the Natural Resources Conservation Service within the US Department
of Agriculture [@usgs_usdanrcs2013]. These provide a spatially-explicit proxy
for specifically human habitats in the proposed study area. There are two
reasons for this. First, the topography of the region is such that the cost of
travel between watersheds is often quite significant. As a consequence,
individuals are expected to spend more time traveling within watersheds than
between them, all else being equal. Second, watersheds are by definition water
sinks, funneling all available run-off into their respective stream networks.
Given the extreme aridity of the proposed project area, this also makes them
critical resource sinks, especially for maize farmers, as watersheds determine
how much water might accumulate at a location, as well as its potential for
irrigation.

While aggregating to the level of the watershed does reduce the spatial
resolution of this analysis, depriving our model of a substantial amount of
environmental variation, several arguments can be made for the simplification.
First, it reduces the computational burdens of this modeling exercise. In
particular, it makes paleoclimate reconstructions more tractable, as we are only
estimating the means within each watershed through time (see below for details).
Second, if there are good theoretical justifications for using certain polygons
as spatially explicit habitats (like the ones outlined above), they can be
relied upon to identify real, presumably causal, relationships in the data.
Third, the aggregated site data can be shared without giving away site
locations, making it easier for others to reproduce the analysis. Finally,
depending on the question being asked, disaggregated site locations may offer
only false precision. This is likely the case for the current analysis as
Fremont were opportunistic hunters [@morgan_etal2012], so many of the sites in
our database are likely temporary logistic hunting sites in proximity of their
actual residential locations, but crucially the available data make it hard to
differentiate these kinds of sites.

## Site Data

To identify Fremont sites in the project area, we rely on site records and
cultural resource reports hosted by the cultural resource information systems in
Nevada (NVCRIS) and Utah (Sego) with permission from the respective State
Historic Preservation Offices. Data collection is part of a larger NSF-funded
project (BCS-1921072) that is examining demographic reconstruction in the
Bonneville Basin and surrounding areas [see @codding_etal2022;
@codding_etal2023; @contreras_codding2023]. Unfortunately, our dataset does not
represent all recorded Fremont sites in the study area, as we did not review all
site forms in a systematic fashion. Some site forms were chosen at random.
Others were selected opportunistically. Still, the resulting dataset contains
2,248 individual Fremont sites, a considerable number representing one of the
largest samples used for this sort of analysis to date.

These site data include everything from small ceramic scatters to massive
Fremont villages like Five Finger Ridge [@janetski1998; @janetski_talbot2000].
For this analysis, we assume that these reflect different levels of population
size and settlement intensity. All else being equal, scatters should follow from
shorter stays by fewer people, villages from longer stays by more people. This
is important because population size correlates with a habitat's suitability
according to the IFD model, and our attempts to reconstruct the ecological and
geographic borders of the Fremont rely heavily on the idea that more suitable
habitats will have larger populations at equilibrium.

To get the architectural feature data, we relied primarily on summaries of well
known and named Fremont villages in [@coltrain_leavitt2002] and cross-referenced
those with original site forms and reports, as well as the work of
[@mooney2014], which attempts to reconstruct poorly reported excavation efforts
along the Wasatch Front. Storage features were not included in this count, only
residential features described as pitstructures, pueblos, roomblocks, or
wickiups. All told, this included 16 Fremont villages and 215 residential
features. In several cases, the estimates of the number of features on a site
vary wildly, so we erred on the side of taking the most conservative estimates.
We then weight the count of sites in each watershed by those estimates. So, a
site with no residential features counts as one site, a site consisting of, say,
four pithouses counts as four unique sites, and a site containing a roomblock
with seven rooms counts as seven sites. This brings the total weighted site
count to 2,447, with minimum and maximum values of 0 and 119, respectively.

@fig-overview shows the log density of Fremont sites across watersheds. The log
is used mainly for visualization purposes as a number of high density watersheds
swamped the variance shown on the untransformed scale. Note that if a watershed
has zero sites in it, we used $log(1e-5)$ as $log(0)$ is undefined. It should be
emphasized that we do not have an independent estimate of the absolute number of
Fremont sites in the area, just an estimate based on a thinned sample, so the
map is more appropriately interpreted as showing relative differences in the
density of recorded sites across watersheds.

## Environmental Covariates

For this analysis, we use both topographic and climatological covariates. The
topographic covariate is cost-distance or travel time to perennial streams
(measured in hours). This provides a coarse grained estimate of water
availability, as well as the potential costs of irrigation. To calculate this,
we first use the R package FedData [@bocinsky2020] to download perennial stream
features from the US National Hydrography Dataset [@usgs2022] and a digital
elevation model (DEM) from the US National Elevation Dataset [@usgs2022a]. We
then apply Campbell's hiking function [@campbell_etal2019] to slope estimates
derived from the DEM. This allows us to estimate travel time between grid cells
in the DEM and then calculate the accumulated cost of travel from each perennial
stream to any grid cell within each watershed. We then aggregate these values to
the watershed level by taking their mean.

Climate covariates include net water-year precipitation (PPT, in millimeters,
mm) from October to September and maize growing degree days (GDD, in Celsius, C)
over the growing season from May to September, as these have a large effect on
maize productivity, thus providing important constraints on the patterns of
settlement for those who relied on maize for subsistence. We caution that GDD is
calculated in such a way as to be insensitive to extreme high and low
temperatures (extreme heat and frost, in effect). It is also conceptually
understood to be a measure of accumulated temperature between the last and first
frost-free days, but in our analysis is calculated using hard start and end
dates of May 1 and September 30, under the assumption that these dates will
roughly coincide with those frost-free days.

::: {#fig-covariates fig-align="center"}
![](../figures/covariates.jpg){width="5in"}

Distribution of covariates across watersheds. These include maize growing degree
days (GDD), annual precipitation (precipitation), and cost-distance to streams
(Streams). Lighter values represent larger values, darker colors smaller values.
:::

We hindcast these climate covariates for each watershed and year over the
Fremont sequence from 1550 to 550 years BP (400 to 1400 CE) using the
Correlation Adjusted corRelation (CAR) method implemented in the R package
paleocar [@bocinsky2019]. This method, in effect, regresses modern climate data
against the width of younger tree rings, then uses older rings to predict past
climate trends [for further details, see @bocinsky_kohler2014;
@bocinsky_etal2016]. In this case, we use interpolated precipitation and
temperature estimates from the \~800 m resolution PRISM dataset
[@prismclimategroup2019] and all tree-ring samples in North America, which were
obtained from the International Tree Ring Database using FedData. For each
watershed, we took the median values of precipitation and GDD over the Fremont
sequence.

@fig-covariates shows the geographic distributions of these covariates over
watersheds. Orographic effects (meaning the effects of mountains and elevation)
are rampant in these data, so it is reasonable to expect that the covariates
will exhibit some multicollinearity. As part of the exploratory phase of this
analysis, we, therefore, perform a series of pair-wise tests of correlation
measured using Pearson's R. We also regress the scaled and centered values (or
z-scores) of maize GDD, precipitation, and cost-distance to streams on elevation
using ordinary least squares (OLS). No doubt, this violates many assumptions of
OLS, particularly the identity and independence of the errors, but explanation
was not our main goal with these simple linear models. We are only trying to get
a rough sense of their general response to changes in elevation, which OLS
offers (see @fig-elevation).

Given that our primary covariates all correlate with elevation (reported below),
readers may wonder why we do not simply use elevation in our main analysis. The
reason for this is two-fold. First, using elevation alone would obscure the
trade-off between water availability and temperature. Second, elevation is not
explanatory of settlement behavior. A maize farmer does not choose a location
for farming *because* it is at, say, 1200 m above sea level. They choose that
location *because* it has the right combination of water availability and
temperature, that is to say, because conditions there are optimal for maize
farming. Were those conditions found at some other elevation, the farmer would
be expected to farm there instead.

## Statistical Modeling and Model Evaluation

For this analysis, the outcome of interest is a count $N$ of archaeological
sites per watershed $i$. Here $N$ is assumed to follow a negative binomial
distribution:$$
N_i \sim NB\,(\lambda_i, r)
$$with $r$ being a dispersion parameter and $\lambda_i$ the expected number of
sites per watershed, $E\,(N_i) = \lambda_i$. To estimate $\lambda_{i}$, we model
it as a log response to the linear predictor:$$
log(\lambda_i) = \alpha + \sum_{k=1}^{K}{s(x_{ik})} + \epsilon_i
$$where $\alpha$ is the intercept, $\epsilon_i$ is the error term, and $s$ is a
smooth function applied to the $K$ covariates - growing degree days,
precipitation, and cost-distance to streams. The smooth is intended to account
for potential non-linear responses to the covariates. In our model, $\epsilon_i$
incorporates an exponential covariance model to account for spatial
autocorrelation in the untransformed residuals. The complete model is fit within
a generalized additive mixed model (GAMM) framework using the R package mgcv
[@wood2004].

Importantly, we add to the model specification a constant offset for the log of
the area of each watershed, in effect making this a model of population
*density*. This is meant to address the idea that larger watersheds will have
more sites just as a matter of chance. While our model as currently specified
includes the intercept, we caution that such an estimate cannot be interpreted
in terms of the abundance of Fremont sites, as that would assume that the
absolute site counts can be inferred from this sample, which is implausible. At
best, the intercept tells us about the abundance of *recorded* sites when
$\sum_{k=1}^{K}{f(x_{ik})}=0$. On that note, we also include a parametric term
for the proportion of each watershed classified as federally protected land.
This is to account for anthropogenic impacts to archaeological resources that
can bias our sample. Notably, for these data, those are almost entirely the
result of Euro-American colonization of the area, in particular, the impacts of
modern farming techniques and the development of the Wasatch Front as the urban
core of the modern state of Utah. This is a coarse metric, but our expectation
is that the greater the share of a watershed with federal protections, the
smaller the effect of these anthropogenic impacts, the more sites we should
observe.

Model evaluation includes checks for concurvity or non-linear correlation in the
smooth terms of the GAMM, as well as a Variance Inflation Factor test on the
parametric or linear terms. We also test for spatial autocorrelation in the
untransformed residuals using Monte Carlo simulations of Moran's I. After
fitting the full model, we found that several smooth terms had effective degrees
of freedom (EDF) equal to one, suggesting no non-linear response in the data.
We, therefore, remove the smooth terms for those covariates in the final model,
leaving only precipitation with a potential non-linear effect.

All analyses are conducted in the R programming language and environment
[@rcoreteam2022]. For details of this analysis, please see the Supplement.

# Results

```{r}
#| label: tbl-results
#| tbl-cap: Model Results
#| output: asis

# tables in quarto are still a hot mess
cat(
"
\\begin{table}[]
\\centering
\\begin{tabular}{lllll}
\\textbf{Parametric Terms}  & \\textbf{exp ($\\beta$)}  & \\textbf{std.er} & \\textbf{t} & \\textbf{p-value} \\\\ \\midrule
Intercept                   & 0.00                      & 1.40             & -7.13       & \\textless{}0.001 \\\\ 
Maize GDD                   & 1.00                      & 0.001            & ~4.15       & \\textless{}0.001 \\\\ 
CD to Streams               & 0.67                      & 0.156            & -2.57       & ~~0.011           \\\\ 
Protected                   & 4.82                      & 0.497            & ~3.16       & ~~0.002           \\\\ \\bottomrule
                            &                           &                  &             &                   \\\\ 
                            &                           &                  &             &                   \\\\ 
\\textbf{Smooth Terms}      & \\textbf{edf}             & \\textbf{ref.df} & \\textbf{F} & \\textbf{p-value} \\\\ \\midrule
s(Precipitation)            & 3.25                      & 3.25             & 11.1        & \\textless{}0.001 \\\\ \\bottomrule
\\end{tabular}
\\end{table}
"
)

```

```{r}

here::here("R", "regression-table.R") |> source()

```

As shown in @tbl-results, all linear (or parametric) coefficients are
significant in the final model. The intercept `r q[["Intercept"]]` is close to
zero when transformed back onto the response scale, likely reflecting the large
number of watersheds with zero Fremont sites, though as mentioned above, the
intercept estimate is only a measure of the recorded site abundance. The
proportion of watersheds falling under federal protections `r q[["Protected"]]`
has a positive effect on site counts, thus confirming our expectation that more
sites occur in watersheds with a larger proportion of federal land. Site counts
increase in watersheds with larger GDD values, though the effect is small
`r q[["Maize GDD"]]`, and site counts decrease with distance from perennial
streams, though again the effect is small `r q[["CD to Streams"]]`. The only
smooth term in the final model is average precipitation over the water-year,
which is significant `r q[["s(Precipitation)"]]`. The effective degrees of
freedom (EDF) for precipitation suggests a strong non-linear effect of that
covariate on site counts, which increases up to about 450 mm and decreases
thereafter.

A concurvity test on the final model was not conducted as there was only one
smooth term. As expected, a VIF test for linear correlation in the parametric
terms of the final model shows some evidence of correlation, particularly for
GDD, which was close to 6, but this value is within acceptable limits at a
moderate threshold. The Moran's I test suggests that there is no spatial
autocorrelation in the untransformed residuals of the final model (I = 0.233,
rank = 381, p = 0.476).

::: {#fig-responses fig-align="center"}
![](../figures/marginal_response.jpg){width="4.5in"}

Partial dependence plots. Each plot shows the response of site counts to changes
in a target covariate while holding all other variables at their mean.
:::

# Discussion

With this analysis, we seek to reconstruct the Fremont maize farming niche by
estimating relative differences in suitability across watersheds and
environmental gradients. Our approach involves the use of a Negative Binomial
GAMM model of the feature weighted counts of archaeological sites, under the
assumption that more suitable watersheds should have more sites at equilibrium.
Results from that model indicate that Fremont maize farmers generally preferred
warmer watersheds with longer growing seasons, moderate levels of precipitation
(in the range of 400-600 mm), and greater access to perennial streams, as shown
by the marginal response plots in @fig-responses.

## The Data Quality Problem in Fremont Research

Fremont research focused on individual, well-dated, and excavated sites can and
often does describe the processes acting on those sites in great detail, but any
attempt to generalize to a whole region runs into a familiar problem, namely
that the Fremont sequence suffers from poor chronological resolution due to a
paucity of direct dates and a more or less undefined ceramic chronological
record. The consequence for this analysis is that we have to use (i) total site
counts in each watershed, not the contemporaneous site counts, and (ii) the
median values of the climate variables over the entire Fremont sequence rather
than over the years that individual Fremont sites were occupied. While there are
radiocarbon dates for Fremont sites in the study area, they represent only a
small fraction of known Fremont sites, so we can use them to get better
chronological control, but only at the expense of losing spatial resolution.

Fortunately, there are good reasons to believe that we have wrestled a
meaningful signal from the noise in these temporally unresolved spatial data.
For one thing, it is unlikely that inter-watershed variability would have
changed substantially over the Fremont sequence. Over the last two thousand
years, wetter watersheds would have tended to be wetter than dryer watersheds,
and warmer watersheds would have tended to be warmer than colder watersheds. The
central tendency of the climate in each watershed also tells us which of them
would, in general, have been better for maize farming, even if we cannot yet
define diachronic variation in watershed suitability. So, the inference is
simply that those typically better watersheds would have attracted maize farmers
more often and, over time, would have accumulated a larger archaeological maize
farming assemblage, too.

## The Relative Contribution of Foraging to Fremont Diets

Many Fremont scholars are convinced that foraging had to contribute a
substantial portion to the diet of Western Fremont populations
[@madsen_simms1998; @simms1999]. If true, this would pose a substantial obstacle
to any inferences about maize suitability based on the distribution of Fremont
archaeological materials, as those inferences would conflate settlement patterns
owing to constraints on foraging and settlement patterns owing to constraints on
farming. This is problematic because it strikes at the heart of the
interpretation we propose for our model. In fact, if true, it would basically
nullify the whole paper, so it is worth confronting the question head-on: to
what extent did foraging contribute to the diet of Fremont maize farming
populations in the eastern Great Basin?

Unfortunately, the kinds of observations that might help us arrive at an
unequivocal answer to this question are frustratingly few and far between. The
most compelling evidence currently available is provided by stable isotope
analysis of a small number of Fremont burials [@coltrain1993;
@coltrain_leavitt2002]. The results of those analyses suggest that the
contribution of maize to Fremont diets was on the order of 75 to 80%, a fraction
comparable to what one would find in the San Juan Basin, where there is no doubt
that maize farming was the primary mode of subsistence during the time period in
question. The one exception to this pattern involves individuals from the GSL
Wetlands, who were at least part time foragers. But those exceptions are limited
to the latter half of the Fremont sequence, after 1100 years BP (850 CE), and
they don't show signs of full reliance on foraging until nearly 800 years BP
(1150 CE). This is consistent with the suggestion that substantial foraging (and
not just opportunistic foraging) would have been relegated mostly to the margins
of areas where maize farming occurs [@allison2008]. In the study area, that
would include areas surrounding and to the north of the GSL, as well as the West
Desert and northern Nevada.

There are two additional things to note here. First, our results have important
parallels with the results reported in [@yaworsky_etal2023], where the Fremont
being studied are unambiguous maize farmers. This suggests that we are also
modeling maize farming populations. The other point to mention is that our model
incorporates an exponential covariance matrix, so if there were any systematic
deviations around the north and northwest periphery of the study area, those
would at least partially be accounted for in the model. In other words, the
primary trend driven by well understood maize farming sites in the core of the
study area (following the I-15 corridor south from Salt Lake) should not be
noticeably affected by those deviations at the periphery.

## Trade-off Between Precipitation and Temperature

While site counts in our model exhibit a positive linear response to maize GDD,
this should not be expected to continue indefinitely for the obvious reason that
maize has an upper threshold temperature at which point it no longer grows.
Plus, increased temperatures trade-off with precipitation, with warmer
conditions being better for plant growth but at the same time increasing water
demand on the plant to keep up with growth [@ramankutty_etal2002]. This
trade-off probably also explains the curious result that site counts do not
increase linearly as a function of precipitation, but instead decline after 500
mm.

Provided that the trade-off between precipitation and temperature is driven
largely by elevation in this region, it would seem likely that the western
Fremont chose site locations at elevations with the best combination of
temperature and precipitation for maize farming in the region, as evidenced by
the estimated distribution of sites across watersheds (see @fig-thresholds),
though we caution that watersheds are coarse grained units of analysis. Given
the topography of the study area, this trade-off would also suggest that areas
suitable for maize agriculture take the form of a thin band around the lower
slopes of mountain ranges.

It is tempting to think of this elevation band as a Goldilocks Zone for maize
agriculture [@yaworsky_etal2023], but we should take care not to over-interpret
this metaphor, for the elevation band is an area of overlap for continuous
ecological gradients, not a region demarcated by anything like a perfect line.
There is also variation in site density within that band, perhaps suggesting
missing covariates in our model. The unaccounted for variation could also be
owing to the fact that what matters most for maize agriculture is not
necessarily where precipitation falls, but where it accumulates once it enters
the stream flow network. Of course, that will largely be a function of elevation
in this region, or changes in elevation, so it will still be in proximity to
this optimal elevation band, though perhaps not perfectly coincident with it.

At any rate, given these considerations, it would perhaps be more accurate to
characterize the trade-off confronting the Fremont as one between what we might
call water availability (including general precipitation across the stream flow
network, water runoff, proximity to perennial surficial sources, soil moisture,
and irrigation potential, among other things) and temperature (including, but
not limited to, the length of the growing season and the number of frost-free
days).

## Water Management and the Rain-fed Maize Farming Niche

Based on previous research [@benson2011; @adams_etal2006; @bellorado2010],
@bocinsky_kohler2014 place minimum thresholds for precipitation and maize GDD at
300 mm and 1000°C GDD, respectively. Importantly, these are thresholds for
dry-farming, which is typically defined as farming free of irrigation
[@benson2011; @cordell_mcbrinn2012; @varien1999], so they can be loosely
interpreted as minimum temperature and precipitation levels required to grow
maize through nothing more than planting and harvesting. The right panel in
@fig-thresholds shows the watersheds whose median values for maize GDD and
precipitation over the Fremont sequence are above those values. This is similar
to Bocinsky's concept of "refugia," but rather than calculate the proportion of
years in niche, we are using median values, so refugia are defined here as
watersheds that spend at least half the Fremont sequence in the rain-fed maize
farming niche ($\geq$ 500 years, in this case).

::: {#fig-thresholds fig-align="center"}
![](../figures/threshold_map.jpg){width="5in"}

The map on the left shows the log density of feature-weighted site counts as
estimated by the model, with lighter colors representing larger relative
densities, darker colors representing smaller relative densities. The map on the
right shows which watersheds are in the rain-fed maize farming niche for at
least half of the Fremont sequence. The light green color indicates those that
are in the niche.
:::

It appears that these models tend to agree on the best watersheds for maize
farming, though there are some notable exceptions. The western slopes of the
Stansbury Range in the north central part of the study area (just west of Salt
Lake City) and parts of the Parowan Valley and surrounding areas in the
southeast part of the study area (the area roughly midway between Cedar City and
Richfield) are estimated by our model to have non-negligible densities, but they
fall outside of the niche. Other areas, like much of the Salt Lake valley appear
to be in the niche, but are estimated by our model to have relatively low
densities. These discrepancies are at least partially explained by limitations
of our methodology. For one thing, including the proportion of federal land as a
covariate is a very crude way of accounting for sampling bias, so our model
might not be picking up on the fact that the Salt Lake valley was actually a
highly suitable area for maize agriculture, as evidence to that effect has
largely been destroyed. The discrepancy may also be explained by the fact that
we have aggregated to the watershed level, thus obscuring important
environmental variation across both the landscape in general and site locations
in particular. Were we to build a model of the disaggregated data, we would
likely find that many agricultural sites do, in fact, fall in the rain-fed maize
agricultural niche, even if the watersheds that contain them do not.

But setting aside those methodological issues, some much more interesting
explanations may be offered, at least from the perspective of theory. Consider
the fact that of the 183 watersheds in the study area, 87% (n=159) have median
GDD values greater than 1000°C, but only 54% (n=98) have median precipitation
values above 300 mm, and a scant 40% (n=74) meet both requirements. This would
suggest that precipitation is the primary limiting factor when it comes to the
choice of watersheds in which to dry farm, for there are many more watersheds
that meet the minimum GDD requirements than meet the minimum precipitation
requirements. One important implication of this is that dry farming would have
been very nearly impossible across most of the study area over the Fremont
sequence and irrigation or some other form of water management more or less
necessary. We note that even within the niche, which we have defined somewhat
liberally as those watersheds that are included for at least half the Fremont
sequence, water management would almost certainly have been a necessity
[@boomgarden_etal2019; @matson_etal1990; @spangler_etal2010]. Of course, the
type and intensity of water management would probably differ between sites, with
more intensive management occurring where there is less precipitation and less
intensive management where there is more precipitation, whether that
precipitation falls in the watersheds themselves or in the upstream watersheds
that feed into them.

## The Spread of Maize Agriculture

Our results comport with previous work suggesting that maize farmers in the
eastern Great Basin preferentially targeted a certain elevation band that
coincides with alluvial fans, forming a rim around the lower slopes of mountains
and ridges on or near important drainages and floodplains. The ruggedness and
aridity of this region led to extreme circumscription, with the costs of
intensive farming outside these areas being so severe as to render that
alternative virtually unsustainable. For maize farmers, the Great Basin should,
thus, have the look and feel of an island biogeography, with mountains rising up
like islands out of an ocean of desert sand and sagebrush.

To account for the timing and tempo of island colonization among food producers
in Oceania, @kennett_etal2006 offer an extension of the IFD framework that might
actually apply to the Great Basin. Their model depends crucially on the
potential for farming populations to introduce economies of scale or Allee
effects, with suitability increasing with increasing population at low
densities. They also assume that individuals will prefer nearer habitats to
those that are farther away, under the assumption that greater distances impose
greater settlement costs, all else being equal. Together, these assumptions
suggest a pulsating pattern of maize spread into the North American Southwest,
with maize agriculture growing, then spreading, growing, then spreading, as
appears to have happened in Oceania.

This need not be a wholesale movement or adoption of agricultural practices,
either. Early on, subsistence would have varied along two dimensions: levels of
residential mobility and levels of cultigen adoption. While some individuals
would have been more mobile, moving from isolated patch to isolated patch,
others would have been more sedentary, typically tethered to a productive
wetland or riparian area. As local populations increase, the benefits of
increased sedentism and cultigen adoption would slowly begin to outweigh the
benefits of a more mobile and foraging-centered diet. According to our model,
individuals moving in this direction would tend to concentrate more of their
time and energy in the maize farming niche. This would then have the effect of
increasing population size within a smaller area, possibly leading to scale
effects, which would invite further sedentism from those in the surrounding
area. Following this logic, the patchy adoption of agriculture makes more sense.
The degree and speed with which it occurs is simply a function of how fast
habitat quality declines and how quickly populations grow.

::: {#fig-model-map fig-align="center"}
![](../figures/model-map.jpg){width="5in"}

Geographic distribution of Fremont sites across watersheds. The left panel shows
the relative density derived from observed site counts, and the right panel
shows the relative density from the estimated site counts.
:::

## The Collapse of Maize Agriculture

After farming these regions successfully for over a thousand years, the Fremont
began an abrupt process of abandonment starting around 700 years BP. The cause
of this event is still a matter of some dispute, though most scholars agree that
climate change drove severe reductions in agricultural productivity
[@benson_etal2007; @spangler_etal2019; @thomson_etal2019;
@thomson_macdonald2020; @finley_etal2020]. We follow @lindsay1986 in thinking
this was likely owing to climatic events that, in effect, severed the connection
between the optimal temperature and precipitation levels that made maize farming
a reliable strategy, pulling higher temperatures further down in elevation and
pushing higher precipitation levels further up.

Lindsay suggests that the end of the Medieval Warm Period may have been
responsible for the initial coupling of these ecological gradients, as it
involved higher temperatures and more growing-season precipitation owing to the
northward intrusion of summer monsoons. This period lasted from roughly 1,000 to
600 years BP [@grayson2011; @graumlich1993], so it does coincide with the height
of the Fremont complex. Given the logic of IFD, this likely led to population
spillovers into less suitable habitats [@codding_etal2022]. However, given the
extreme circumscription we have highlighted in the Great Basin, the available
alternatives would have been limited, potentially leading to increasing
population packing across all habitats during this time, thus making the Fremont
especially sensitive to climatic downturns.

The well-documented megadrought at the end of the thirteenth century
[@cook_etal2004] would have been catastrophic in this context, reducing
precipitation levels across the region [@benson_etal2007], but also leading to
reductions in water accumulation within the streamflow network. So, just as
temperatures are beginning to cool, precipitation levels collapse. By extension,
this would have decreased overall water availability and, thus, the potential
for irrigation. Opportunities for the Fremont to recover from such a drought
would have been limited, too, as the end of the Medieval Warm Period was also
the beginning of the Little Ice Age, a time during which conditions were
generally colder and wetter [@mann_etal2009], further pulling apart the
combination of temperature and precipitation upon which the Fremont so long
depended.

\pagebreak

# References

::: {#refs}
:::